Ideal model selection is impossible

### 2014-10-30 Facticity/Facticity/paper.tex
...

\newcommand{\M}{\mathcal M}
\newcommand{\K}{\mathcal K}
\newcommand{\X}{\mathcal X}
\newcommand{\C}{\mathcal C}
\newcommand{\B}{\mathbb B}
\newcommand{\D}{\Delta}
\newcommand{\N}{\mathbb N}
\newcommand{\tn}[1]{\textnormal{#1}}
\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
\newcommand{\concat}{\oplus}
\newcommand{\symb}[1]{\texttt{#1}}
\newcommand{\p}{\quad\text{.}}
\newcommand{\tuple}[1]{\left\langle{#1}\right\rangle}

@@

@@

\begin{definition}
  Given a model $f$ and model class $\M$, define model-based complexities by
  \begin{align*}
  C^f(x) &= \min\{|y|:f(y)=x\}\\
  L^f(x) &= K(f)+C^f(x)\\
  L^{\M}(x) &= \min\{L^f(x)\mid f\in\M\}
  \end{align*}
\end{definition}

\hrule

\begin{lemma}
  Let $\M$ be any model class and let $\M'\subseteq \M$ be a sub-model with the following properties:
\begin{enumerate}
  \item The function $g(x):=\argmin_{f\in\M'}L^f(x)$ is computable and uniquely defined for all $x$
  \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
\end{enumerate}
Then there is an enumeration $\phi_1,\phi_2,\ldots$ such that $F^{\M,\phi}(x)=K^\phi(g(x))+1$ for all $x$.
\end{lemma}
\begin{proof}

Construct $\phi_1,\phi_2,\ldots$ such that
\begin{enumerate}
\item For all $f\in\M'$ have $K^\phi(f) = K^\psi(f)+1$ (zelf bewijzen)
\item For all $f\not\in\M'$ have $K^\phi(f)>K^\psi(f)+c$.
\end{enumerate}

Let $\psi_1,\psi_2,\ldots$ be an acceptable numbering of the partial recursive functions, and let $g'(x)=i$ such that $\psi_i=g(x)$. Now define a new ordering $\phi_1,\phi_2,\ldots$ by
\[
\begin{cases}
\phi_{0i}(y)&=\psi_i(y),\qquad\tn{if $i=g'(x)$ for some $x$, and $\infty$ otherwise}\\
\phi_{1^r0i}(y)&=\psi_i(y)\\
\phi_{\tn{all-ones}}(y)&=\infty
\end{cases}
\]

Voor alle $x$ kiezen we een makkelijk te vinden, en goedkope representatie met input $y$. Die maken we extra goedkoop in de nieuwe enumeratie. Dan wordt ie altijd gekozen.
\end{proof}
\hrule
\begin{lemma}
  Let $\M$ be a model class with the property that there is a constant $c$ and a computable function $g$ such that for all $x$ we have $\gamma(x\mid g(x))\le c$.
  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $K^{\M,\phi}(x)-F^{\M,\phi}(x)$ is bounded.
\end{lemma}
\begin{proof}

>>>>

\newcommand{\M}{\mathcal M}
\newcommand{\K}{\mathcal K}
\newcommand{\X}{\mathcal X}
\newcommand{\C}{\mathcal C}
\newcommand{\B}{\mathbb B}
\newcommand{\D}{\Delta}
\newcommand{\N}{\mathbb N}
\newcommand{\tn}[1]{\textnormal{#1}}
\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
\newcommand{\concat}{\oplus}
\newcommand{\symb}[1]{\texttt{#1}}

\newcommand{\p}{\quad\text{.}}

\newcommand{\tuple}[1]{\left\langle{#1}\right\rangle}

@@

\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings. There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
\begin{align*}
K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
= K^\psi(f)
\end{align*}
We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
\end{proof}

@@

\begin{definition}
We call $\phi_1, \phi_2, \ldots$ a \emph{partial computable numbering} of some subset of the partial computable functions, if there exists a partial computable function $u: \N \times \B \to \B \cup \infty$ such that $u(i, p) = \phi_i(p)$. We call $u$ the universal model for the numbering. We call a numbering \emph{optimal} if there is a constant $c$ such that for all $i$, we have $|i| \leq K(\phi_i) + c$.
\end{definition}

\begin{lemma}
There exists an acceptable, optimal partial computable numbering of the partial computable functions.
\label{lemma:optimal-numbering}
\end{lemma}
\begin{proof}
Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering, with some $\psi_{\text{div}}$ with is undefined for all its arguments. Let 
\begin{align*}
	\phi_{\bar\imath p} &= \psi_{\psi_i(p)} \qquad\text{if $\psi_i(p)$ is defined}\\
	\phi_{\bar\imath p} &= \psi_{\text{div}} \qquad\text{otherwise}\\
	\phi_{j} &= \psi_{\text{div}} \qquad\text{if $j$ can not be read as $\bar\imath p$}
\end{align*}

\noindent To show that this numbering is acceptable. We must provide two computable function to translate t and from a given acceptable numbering. We use $\psi$ for this purpose. To find a $\phi_j$ which corresponds to a given $\psi_i$, we can simply use the identity function $\psi_d(x) = x$. For a given $\psi_i$, the function $\phi_{\overline{d}i}$ computes the same function.

To translate from $\phi_j$ to $\phi_i$, we are faced with a problem. If we are given $j = \bar\imath p$, we cannot decide algorithmically whether $\psi_i(p)$ is defined. However, we can construct the function $\psi_r(\bar\imath p, q)$ such that it computes $\psi_r{\psi_i(p)}(q)$ if $\psi_i(p)$ is defined and is undefined otherwise, or if the first argument cannot be read as $\bar\imath p$. For a given $j$ we can also find an $r'$ such that $\psi_{r'}(q) = \psi_(j, q)$. Thus, given any $\phi_i$ we can construct and return this function.

Finally, we'll show that the minimum length of an index for function $f$ in $\{\phi_i\}$ is equal to $K(f)$ up to a constant:
\begin{align*}
K^{\psi}(f) &= \min\left\{K^{\psi}(j)\,:\,\psi_j = f \right\} \\
&= \min_j\left\{\min_{i, p} \{|\bar\imath p| \,:\,\psi_i(p) = j\} \,:\,\psi_j = f\right\} \\
&= \min_{i, p} \left\{|\bar\imath p| \,:\,\psi_i(p) = j, \psi_j = f\right\} \\
&= \min_j \left\{|j|\,:\, \phi_j = f \right\}
\end{align*}
\end{proof}

\begin{lemma}
Let $M$ be the class of \emph{singleton functions}, that is, the functions which produce a single output for all arguments, including the function which is undefined for all its arguments. There exists an optimal partial computable numbering of $M$.
\end{lemma}
\begin{proof}
Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Let $\phi_{\bar\imath p}(\cdot) = \psi_i(p)$ \ldots. 
\end{proof}

\begin{definition}
  Given a model $f$ and model class $\M$, define model-based complexities by
  \begin{align*}
  C^f(x) &= \min\{|y|:f(y)=x\} \text{,}\\
  L^f(x) &= K(f)+C^f(x) \text{,}\\
  L^{\M}(x) &= \min\{L^f(x)\mid f\in\M\}\text{.}
  \end{align*}
\end{definition}

\begin{lemma}
  Let $\M$ be any model class and let $\M'\subseteq \M$ be a sub-model with the following properties:
\begin{enumerate}
  \item There exists an optimal partial computable numbering $\eta_1, \eta_2, \ldots$ for $M'$,
  \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
\end{enumerate}
There is an enumeration $\phi_1,\phi_2,\ldots$ of $\M$, such that $S^\phi_{\M}(x) = S^\eta_{\M'}(x)+1$ for all $x$.
\end{lemma}
\begin{proof}
Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Let $u$ be the universal model for the numbering $\eta_1, \eta_2, \ldots$. We define the numbering $\phi$ as follows:
\begin{align}
\phi_{0}i(p) &= u(i, p) \\
\phi1^r0i(p) &= \psi_i(p) \\
\phi_j(\cdot) &= \infty \quad\text{if $j$ contains no zeroes.}
\end{align}
Note that with this definition, we have $K^\phi(f) \leq K^\eta(f) \leq K^\psi(f) + c$ if $f \in \M'$. If $f$ is not in $\M'$, we have $K^\phi(f) = K^\psi(f) + r + 1 - c'$, by the same argument as used in the proof of Lemma~\ref{lemma:building-block}.

Let $x$ be some finite string, and let $(i, p)$ be the pair such that $\psi_i(p)$ witnesses $L^M_\psi(x)$. By assumption, $L^{\M'}_\psi \leq L^M(x) +c$. Thus, we have a pair $(j, q)$, with $\eta_j(q) = x$ and $K^\eta(j) + |q| \leq K^\phi(i) + |p| + c$. If we choose $r = c' + c+ 1$, we know that $\phi$ will have representations $(1^r0i, q)$ with code length $L^M_\psi(x) + c+ 1 $, and $(0j, q)$ with code length less than or equal too $L^{\M}_\psi(x) +c$. This ensures that the sophistication under $\phi$ will be $K^\phi(\eta_j) = K^\eta(\eta_j) + O(1) = F^{\M'}_\eta(j)$.
\end{proof}

\begin{lemma}
  Let $\M$ be any model class and let $\M'\subseteq \M$ be a sub-model with the following properties:
\begin{enumerate}
  \item There exists a (non-optimal) computable numbering $\psi_1, \psi_2, \ldots$ for $M'$,
  \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
\end{enumerate}
Then there is an enumeration $\phi_1,\phi_2,\ldots$ such that $|F^{\psi}_{\M'} - \leq F^{\phi}_{\M}| \leq c$ for infinitely many $x$.
\end{lemma}
\begin{proof}

\hrule
\begin{lemma}
  Let $\M$ be a model class with the property that there is a constant $c$ and a computable function $g$ such that for all $x$ we have $\gamma(x\mid g(x))\le c$.
  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $K^{\M,\phi}(x)-F^{\M,\phi}(x)$ is bounded.
\end{lemma}
\begin{proof}
